\chapter{Conclusion}
\hrule
\vspace{40pt}

\section{Summary}
In this paper we explored various machine learning models for predicting high frequency returns
for four of the main cryptocurrency futures trading pairs, BTCUSDT, ETHUSDT, MATICUSDT and SOLUSDT. 
We constructed a novel data set from a live WebSocket stream from Binance, the worlds largest centralized cryptocurrency exchange.
Following \cite{CONT2013}, we defined a stationary transformation of the raw orderbook data called orderflow, 
and motivated its definition by showing that under a stylized model of the orderbook, orderflow can be used to measure linear price impact
of orderbook events. We then defined our modelling objective as a three-class classification problem, predicting the direction of the smoothed
mid-price returns for several prediction horizons. For this task, we introduced several traditional machine learning and also deep learning
models that use either the orderflow representation or the raw orderbook representation. Specifically, from traditional ML we introduced
Logistic Regression and XGBoost models. For our deep learning models, we used the deepLOB model from \cite{ZHANG2019} and the deepOF model
from \cite{KOLM2023} which use the raw orderbook/orderflow representations respectively.
We trained, validated and tested our models for each trading pair and multiple prediction horizons on 8 windows from 05-02-2024 to 25-02-2024.
We presented the results on the test set and also evaluated the feature importances for each XGBoost model.

\section{Conclusions}
From our results section we showed that we can predict the direction of smoothed mid-price returns at 
multiple short term time horizons for our Binance WebSocket data with good performance.
We also saw that the price features are not significantly used and we concluded that the orderflow representation is far superior
to the raw orderbook data representation for this task. This agrees with the findings of \cite{KOLM2023}.
We found that simpler traditional machine learning models have very competitive performance when compared with the much larger
and more complex deep learning models. This is a significant result, since it challenges the idea that we need a large and complicated
deep learning model to achieve state of the art performance for orderbook prediction. \cite{ZHANG2019}, \cite{KOLM2023}, \cite{LUCCHESE2024}.
This is very good news for practical application, since simpler models are much more explainable, which is a very important quality
for many quantitative applications.

\section{Future Work}
As previously mentioned, it would be interesting to see how our models compare with those found in \cite{ZHANG2019}, \cite{KOLM2023}, \cite{LUCCHESE2024} when applied to their data.

Also, in this paper we only explored 4 trading pairs, when in fact Binance has over two-hundred perpetual futures trading pairs available.
It would be interesting to apply the same methodology presented in this paper to a larger group of trading pairs and explore
the differences across pairs.

A major advantage of using the Binance WebSocket data is that it is freely available and live at any time. Our next steps
would be to adapt the models presented here to be, \textit{online}, i.e. continously learning and adapting in real time.
Then we could continuously test the models in real time for a longer period of time. This would show definitively
whether this methodology is practical for real world trading applications.



